\thema

### 2016-03-03 modules/6_sec_infrastructure.tex
Continued 6_sec_infrastructure. Minor additions to chapter structure in that chapter.

\chapter{Docker in Unternehmen/Clound-Infrastrukturen}
\label{secInfrastructure}
  Docker wird von den meisten Cloud-Anbietern unterstützt.

  Google: Google Container Engine mit Docker Containern. Mit Kubernetes-Schicht dabei. --> komplette FOSS infrastruktur bei google
  Amazon: Elastic Bean Stack (fuer web applications) und auch EC2-Container als EC2-Container-Service (ECS)

  \section{Microsoft Azure}
    % https://docs.docker.com/v1.9/engine/installation/azure/

@@

  \section{IBM SoftLayer}
    % https://www.docker.com/IBM
    % https://docs.docker.com/v1.9/engine/installation/softlayer/
    % Treiber: https://docs.docker.com/machine/drivers/soft-layer/
    % http://blog.softlayer.com/tag/docker

@@

  % -------

@@

  % Deploying Docker on your infrastructure:
  % ^   \cite[S.6+7]{dockerSecIntro}

  % Ganz gutes Bild des Aufbaus in virtVSContainer_2014.... s.2 .
  % Auch text in linker spalte

  % Gutes Intro von Docker mit public cloud orechestration:
  % ^   \cite[S.4+5]{virtVSContainer}

>>>>

\chapter{Docker in Cloud-Infrastrukturen}
\label{secInfrastructure}
  Wie in der Einleitung gezeigt, bietet der Einsatz von Virtualisierungslösungen, insbesondere Container, in Rechenzentren einige Vorteile.
  % Neben der technischen Überlegenheit, hat es Docker durch ein sehr erfolgreiches Marketing geschafft, als Heilmittel vieler Probleme in IT-Abteilungen in Unternehmen geschätzt zu werden.

  Durch den hohen Bekanntheitsgrad und der Popularität von Docker in Entwickler- und Managementkreisen sind Anbieter von Public-Clouds gezwungen, eine Unterstützung von Docker in ihr eigenes Portfolio aufzunehmen. Tatsächlich reagieren die großen Anbieter wie \emph{Amazon}, \emph{Microsoft}, \emph{IBM} und \emph{Google} mit vielseitigen, häufig Docker-basierten Containerangeboten. Die etablierten Cloud-Ausprägungen SaaS, PaaS und IaaS werden durch explizite Containerangebote ergänzt, sodass zur Zeit der Begriff CaaS (Container as a Service) als weitere Angebotsform von Cloud-Angeboten Gestalt annimmt. Die Auswirkungen von Docker auf genannte Anbieter sind in Kapitel \ref{publicCloud} beschrieben.

  Auch in Private Clouds ist Docker ein Thema. Die in Kapitel \ref{secLinux} und \ref{secEcosystem} aufgeführten Eigenschaften von Docker eignen sich für eine unternehmensinterne Anwendung der Technologie, wie in Kapitel \ref{privateCloud} näher erläutert wird.

  %Google: Google Container Engine mit Docker Containern. Mit Kubernetes-Schicht dabei. --> komplette FOSS infrastruktur bei google
  %Amazon: Elastic Bean Stack (fuer web applications) und auch EC2-Container als EC2-Container-Service (ECS)

  \section{Public Cloud}
  \label{publicCloud}
    Während der Begriff CaaS von Unternehmen hauptsächlich für Marketingzwecke genutzt wird, sind sich die großen Vertreter auf dem Virtualisierungsmarkt selbst nicht über die Bedeutung des Begriffs einig. \emph{Google} z.B. versteht darunter eine Mischform aus PaaS und IaaS, die von ihrer eigenen Entwicklung \emph{Kubernetes} ermöglicht wird \cite[S.12]{http://www.slideshare.net/brendandburns/defrag-2014-41815642?ref=http://thenewstack.io/google-offers-container-as-a-service-to-define-kubernetes-place-in-the-cloud-economy/}.
    % TODO: Kann man hier sagen "nach Aussage von Google"? Weil die Aussage eigtl. von Vortrag von Brendan Burns kommt, also einem Google-Mitarbeiter.
    \emph{Docker} hingegen umschreibt bei der eigenen Definition von CaaS das Konzept von DevOps, das einen von Docker ermöglichten, flexiblen Workflow der Zukunft realisiert. \glqq{}Der Weg zu CaaS ist mit Docker gepflastert \grqq{} ist eine Aussage eines Artikels des offiziellen Docker-Blogs \cite{https://blog.docker.com/2016/02/containers-as-a-service-caas/}.

    So versuchen Unternehmen dem Begriff CaaS nach eigenen Vorstellungen Gestalt zu verleihen und ihn zum eigenen Vorteil zu nutzen.

    Fest steht, dass Docker als disruptive Technologie von allen Anbieter von Public Clouds akzeptiert und unterstützt werden muss. Diese Tatsache beruht weniger auf der technischen Überlegenheit von Docker gegenüber anderen Containerlösungen oder der Hypervisor-basierten Virtualisierung, sondern auf einem einfachen, automatisierbaren Workflow, der im Modell von Continuous Integration und Continuous Delivery von Docker ermöglicht wird. Die Innovation, mit der sich IT-Unternehmen aktuell konfrontiert sehen, geschieht dadurch vorrangig auf Prozess- und Geschäftsebene.

    Wie große Public Cloud-Anbieter Docker in ihr bestehendes Produktportfolio integrieren und welche Sicherheitsfeatures diese anbieten, ist am Beispiel von \emph{Azure} und \emph{SoftLayer} in den nächsten Abschnitten aufgezeigt.

    \subsection{Beispiel: Microsoft Azure}
      % https://docs.docker.com/v1.9/engine/installation/azure/

@@

    \subsection{Beispiel: IBM SoftLayer}

      % https://www.docker.com/IBM
      % https://docs.docker.com/v1.9/engine/installation/softlayer/
      % Treiber: https://docs.docker.com/machine/drivers/soft-layer/
      % http://blog.softlayer.com/tag/docker

@@

  \section{Private Cloud}
  \label{privateCloud}

@@

  % Deploying Docker on your infrastructure:
  % --------------------------------------------------
  % Both containers and VMs provide isolated environments for running applications on a shared host but from different technical perspectives and can be successfully used separately or together depending on the needs of the application environment.
  % Virtual machines have a full OS, including its own memory management and virtual device drivers. Isolation is provided at the virtual machine level with resources being emulated for the guest OS, allowing for one or more parallel (and, eventually, different) OS to run on a single host. VMs provide a barrier between application processes and bare-metal systems. The hypervisor denies a VM from executing instructions which could compromise the integrity of the host platform. Protecting the host relies upon providing a safe virtual hardware environment for which to run an OS. This architecture has different implications for host resource utilization, but allows for applications with different OS’s to run on a single host.
  % Docker containers share a single host OS across all of the application containers running on that same host. Isolation is provided on a per application level by the Docker engine. Using containers reduces the overhead used per application because the multiple OS instances are avoided. This makes containers lighter weight, faster and easier to scale up or down and can gain higher density levels than full VMs. This approach is only possible for applications that share a common OS, like Linux is used for distributed applications.
  % Regular VMs function in a way that does not allow them to be efficiently scaled down to the level of running a single application service. A VM can support a relatively rich set of applications but running multiple microservices in a single VM without containers creates conflict issues while running one microservice per VM may not be financially feasible for some organizations. Deploying Docker containers in conjunction with VMs allows an entire group of services to be isolated from each other and then grouped inside of a virtual machine. This approach increases security by introducing two layers, containers and VMs, to the distributed application. Additionally this method employs a more efficient use of resources and can increase the density of containers while decrease the number of VMs required for the defined isolation and security goals.
  % Therefore stronger application isolation can be achieved by combining virtualization and containerization than is cost and resource efficient with virtualization alone. Docker containers pair well with virtualization technologies, by protecting the virtual machine itself, and providing defense-in-depth for the host.

  % Running Containers on BareßMetalÖ
  % Containers provide a layer of protection between the host and its applications, isolating between application and the host. This makes it safer to deploy applications on bare-metal when compared to not using any virtualization or container technology. With containers, many application services can be deployed on a single host enabling organizations to gain higher levels of resource utilization out of their infrastructure.
  % However, bare metal deployment does not provide ring-1 hardware isolation, given that it cannot take full advantage of Intel’s VT-d and VT-x technologies. In this scenario containerization is not a complete replacement of virtualization for host isolation levels.
  % Containerization does provide isolation for running applications on bare-metal, which protects the machine from a large array of threats and is sufficient for a wide range of use cases. Users in the following scenarios may not be good candidates to use VMs and can instead use containers; performance-critical applications running on a single-tenant private cloud, where cross-tenant or cross-application attacks are not as much of a concern; or they are using specialized hardware which cannot be passed through to a VM, or which hardware that offers direct-memory-access, thus nullifying the isolation benefits of virtualization. Many users of GPU computing are in this position.
  % Docker containers running on bare-metal have the same high-level restrictions applied to them as they would if running on virtual machines. In neither case, would a container normally be allowed to modify devices or hardware, either physical or virtual

  % Aus CONCLUSION:
  % Containers and Virtual Machines (VMs) can be deployed together to provide additional layers of isolation and security for selected services.

  % ^   \cite[S.6+7]{dockerSecIntro}

  % PaaS Use Case
  % Platform as a Service focuses on providing Language Runtime and Services to developers leaving Infrastructure Provisioning and Management to the underlying Layer. PaaS could use Host OS, VM or Container as hosting Environment for applications. PaaS can be implemented using one of the three scenarios.
  % • Option 1 : Each App runs in an Application level container directly on Host OS where PaaS layer runs
  % • Option 2 : Each App runs in a VM on PaaS : Each App runs in its own VM
  % • Option 3 : Each App runs in a container which shares a common VM or VMs for hosting Apps
  % Option 3 is one of the common implementation choice for most of the PaaS vendors as a combination of the VM and container meets some of the requirements of Isolation, Lifecycle management, fair share of CPU time.

  % Using VMs in PaaS:
  % Paas has an important use for VMs to abstract out the Runtime OS from the underlying hardware. Need for each application to have its own VM is the point of contention. Since containers tend to be much lighter in memory footprint, they are preferred architecture as compared to a VM for each application. VM is are used to host the containers. Other Services in Paas like Authentication, Routing, Cloud Controller, Persistent Storage is still run directly on Virtual Machines.

  % From CONCLUSION:
  % Only few PaaS implementations notably the ones which are open source or are new are using containers as opposed to older ones like GAE, Azure which do not use this concept. Containers have a bright future specially in the PaaS use case provided there is more standardization and abstraction from the underlying kernel and Host OS.

  % mittelmaessiges Bild auf S.2

  %     ^   \cite[S.2+5]{virtVSContainer}
